---
title: "p8130_FinalProject"
author: "Jiaqi Li"
date: "12/9/2017"
output:
  html_document: default
  pdf_document: default
---

```{r load_libraris, message=FALSE, warning=FALSE}
library(car)
library(caret)
library(leaps)
library(psych)
library(boot)
library(janitor)
library(knitr)
library(corrplot)
library(tidyverse)
library(dplyr)

```

```{r read_data, message=FALSE, warning=FALSE, cache = T}

ghc <- readxl::read_excel("GHProject_Dataset.xlsx") %>% 
  clean_names()

ghc$date <- as.POSIXct(ghc$admitdtm, format = "%A, %B %d, %Y")

index.final <- 0

for (i in unique(ghc$patientid)) {
  
  index <- which(ghc$patientid == i)
  final.index <- index[which(ghc$date[index] == min(ghc$date[index]))]
  index.final <- c(index.final, final.index)
  
}

ghc <- ghc[index.final, ]

```

Each VisitID represents a unique visit. However, it is possible that a patient visited the hospital more than once. Summarize the number of visits per patient and if multiple visits per patient, select the first visit (by date) recorded. 

```{r clean_data, message=FALSE, warning=FALSE, cache = T}

factor_var <- c("is30dayreadmit", "mews", "cindex", "icu_flag", "evisit", "gender", "race", "religion", "maritalstatus", "insurancetype")
ghc[, factor_var] <- lapply(ghc[, factor_var], factor) ## convert to factor variables
 

```



```{r continous_table, message=FALSE, warning=FALSE, cache = T}

## descriptive summary for continuous data

ghc_conti_summary <- ghc %>% 
  describe() %>% 
  .[-c(1, 2, 5:10, 12:18,26,27), -c(1, 2, 6, 7, 11, 12, 13)] 

ghc_conti <- ghc[, -c(1, 2, 5:10, 12:18,26,27)]

slr <- function(variable) {
  lm(loshours ~ variable, data = ghc_conti)

}

slr_conti <- map(ghc_conti, slr) %>% 
  map_df(., broom::tidy) %>% 
  filter(term == "variable") %>% 
  select(estimate, p.value) %>% 
  cbind(ghc_conti_summary, .)

CI <- function(variable) {
  a <- lm(losdays2 ~ variable, data = ghc_conti)
  confint(a)
}

CI_conti <- map_df(ghc_conti, CI) %>% 
  .[c(2, 4),] %>% 
  t() %>% 
  cbind(slr_conti, .) 

colnames(CI_conti)[c(7, 9, 10)] <- c("beta1", "CI lower", "CI upper")

conti_summary <- round(CI_conti, 4) %>% 
  kable()

conti_summary
```

```{r categorical_table, message=FALSE, warning=FALSE, cache = T}

## descriptive summary for categorical data

colnames(ghc)[c(5:10, 12:18 )] ## These are categorical variables

ghc_cate <- ghc[,c(5, 7:10, 13:18 )] 
map(ghc_cate, table)
 

  # t() %>% 
  cbind() %>% 
  View()
  
  ## categorical variables frequency table

```


```{r cleaning_table }

## The smallest ever recorded BMI is 7 
## Temperatures higher than 46 or 21 indicate death you are dying.

ghc <- ghc %>%
        filter(temperature < 46,
               temperature > 21,
               bmi > 10,
               o2sat < 100)

```

```{r exploratory_plots, message=FALSE, warning=FALSE, eval = TRUE, cache = T}

ghc_conti <- ghc[, -c(1, 2, 5,6,8, 9, 10, 12:18, 26, 27)] %>% 
  drop_na()  

ghc_conti$mews <- as.numeric(as.character(ghc_conti$mews))
corrplot(cor(ghc_conti[c("mews","bpsystolic","o2sat","temperature","heartrate",
          "respirationrate","bpdiastolic","bmi")]), method = "number")

panel.hist <- function(x, ...){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE, 8)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...){
  usr <- par("usr");on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if (missing(cex.cor)) cex.cor <- 1.2/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor*sqrt(sqrt(r))*5)
}

my_line <- function(x,y,...){
  points(x,y,cex = 6,...)
  abline(a = lm(y~x)$coefficients[1] , b = lm(y~x)$coefficients[2] , ...)
}

pairs(ghc_conti, lower.panel = panel.smooth,
      cex = .5,
      upper.panel = panel.cor,
      pch = 1, bg = "light blue",
      diag.panel = panel.hist, cex.labels = 1, font.labels = 0.2)

## We can see from the panel that the correlation between loshours and losdays2 is 1. It means they are exactly the same thing, which makes sense. So I drop hosdays2

ghc_conti <- ghc_conti %>%
              mutate(log_days = log(losdays2)) %>%
              select(-loshours, -losdays2)
              

pairs(ghc_conti, lower.panel = panel.smooth,
      cex = .5,
      upper.panel = panel.cor,
      pch = 1, bg = "light blue",
      diag.panel = panel.hist, cex.labels = 1, font.labels = 0.2)

```

```{r transformation, message=FALSE, warning=FALSE, cache = T} 
par(mfrow = c(1, 2))

hist(ghc$bpsystolic)
hist(log(ghc$bpsystolic))

hist(ghc$bmi)
hist(log(ghc$bmi))

hist(ghc$bpsystolic)
hist(log(ghc$bpsystolic))

hist(ghc$heartrate)
hist(log(ghc$heartrate))

hist(ghc$respirationrate)
hist(log(ghc$respirationrate)) ## still skewed after transformation

hist(ghc$bpdiastolic)
hist(log(ghc$bpdiastolic)) ## transformation may not need?

hist(ghc$losdays2)
hist(log(ghc$losdays2)) ## Transformation helps
```

Besides these, I also tried several transformations with _evisit_, but none of them looks good.

```{r transformation2, cache = T}

## There a small quantity for these two hospitals

ghc$facilityname[which(ghc$facilityname == "Lenox Hill Hospital" | 
                       ghc$facilityname == "Syosset Hospital")] <- "Other"

## Combining religions

ghc$religion[which(ghc$religion == "Angelican")] <- "Christian"
ghc$religion[which(ghc$religion == "Non Denominational")] <- "Christian"
ghc$religion[which(ghc$religion == "Mormon")] <- "Christian"
ghc$religion[which(ghc$religion == "Hebrew")] <- "Jewish"
ghc$religion[which(ghc$religion == "Angelican")] <- "Christian"
ghc$religion[which(ghc$religion == "Angelican")] <- "Christian"

## Combining race

ghc$race[which(ghc$race == "Native Amer/Alaskan" | 
               ghc$race == "Natv Hawaii/Pacf Isl" |
               ghc$race == "Other/Multiracial"   )] <- "Other"


# There's a couple civil union so marking them within married

ghc$maritalstatus[which(ghc$maritalstatus == "Civil Union")] <- "Married"


# Get month out
# take the log of los days
# Got rid of BMI since there's so many missing values and it's not highly correlated with outcome

ghc_final <- ghc %>% separate(admitdtm, into = c("remove", "Month"), sep = ", ") %>%
               separate(Month, into = c("Month", "remove2"), sep = " ") %>%
               dplyr::select(-bmi, -remove, -remove2, -loshours) %>%
               mutate_if(is.character,as.factor) %>%
               mutate(log_losday = log(losdays2)) %>%
               drop_na()

### Month information

## This is highly significant, there's clearly a bias across months
oneway.test(ghc_final$losdays2 ~ ghc_final$Month)

table <- describeBy(ghc_final$losdays2, group = ghc_final$Month,
                    mat = TRUE)

# Error bars represent standard error of the mean
ggplot(table, aes(x = group1, y = mean)) + 
    geom_bar(position = position_dodge(), stat = "identity") +
    geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se),
                  width = .2,                    # Width of the error bars
                  position = position_dodge(.9))

### The observations for december are from December 29th and 30th.
### The obvious explanation is that this is right censored so we need to get rid of them

ghc_final <- ghc_final %>% filter(Month != "December")

oneway.test(ghc_final$losdays2 ~ ghc_final$Month)

## MEWS Information

## This is highly significant 
## Pooled mews according to the groups in the paper

ghc_final$mews <- as.character(ghc_final$mews)
ghc_final <- ghc_final %>% mutate(group_mews = "ActionRequired")
ghc_final$group_mews[which(ghc_final$mews == "0" | 
                     ghc_final$mews == "1")] <- "Normal"

ghc_final$group_mews[which(ghc_final$mews == "2" | 
                     ghc_final$mews == "3")] <- "Caution"

ghc_final$group_mews[which(ghc_final$mews == "4" | 
                     ghc_final$mews == "5")] <- "Deterioration"

ghc_final <- ghc_final %>%
              mutate(group_mews = as.factor(group_mews),
                     group_mews = forcats::fct_relevel(group_mews, "Normal", 
                                                       "Caution","Deterioration",
                                                       "ActionRequired")) %>%
              dplyr::select(-mews)

oneway.test(ghc_final$losdays2 ~ ghc_final$group_mews)

table <- describeBy(ghc_final$losdays2, group = ghc_final$group_mews,
                    mat = TRUE)

## Questionable wether to include or not since it overlaps with vital signs

# Error bars represent standard error of the mean
ggplot(table, aes(x = group1, y = mean)) + 
    geom_bar(position = position_dodge(), stat = "identity") +
    geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se),
                  width = .2,                    # Width of the error bars
                  position = position_dodge(.9))

ghc_final <- ghc_final %>% 
  mutate(cindex_cat = "severe")


ghc_final$cindex_cat[which(ghc_final$cindex == 0)] <- "Normal"
ghc_final$cindex_cat[which(ghc_final$cindex == 1 | 
                              ghc_final$cindex == 2 )] <- "Mild"
ghc_final$cindex_cat[which(ghc_final$cindex == 3 | 
                              ghc_final$cindex == 4 )] <- "Moderate"

ghc_final <- ghc_final %>% 
                filter(Month != "December",
                       icu_flag == 0 ) %>%
                dplyr::select(-patientid, -visitid, -postalcode, 
                       -facilityzip, -date, -losdays2, -icu_flag, -cindex)



write_csv(ghc_final, "./BMI_FinalProject_cleandata_SAS.csv")

```


```{r models}

MainE_Model <- lm(formula = log_losday ~ is30dayreadmit + cindex_cat + evisit + 
                    ageyear + gender + maritalstatus + facilityname + 
                    insurancetype + bpsystolic + temperature + heartrate + o2sat +
                    respirationrate + bpdiastolic, data = ghc_final)


## Criterion selection based on AIC
step(MainE_Model, direction = "both")

## These are all the remaining criterions
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
} 

## All of them tell you to use the full model with the exception of BIC which inidates a model of 8 variables. 

best(MainE_Model, nvmax = 14)

## This is the final model
MainE_Model <- lm(formula = log_losday ~ is30dayreadmit + cindex_cat + evisit + 
    ageyear + gender + maritalstatus + facilityname + insurancetype + 
    bpsystolic + temperature + heartrate + o2sat + respirationrate + 
    bpdiastolic, data = ghc_final)

par(mfrow = c(2, 2))
plot(MainE_Model)

t <- MainE_Model %>% broom::tidy() %>% kable(digits = 3)

## This gives you the pvalues of testing the model with and without each variable
drop1(MainE_Model, test = "F") %>% broom::tidy() %>% kable(digits = 3)

```


```{r influence_vif, cache = T}
t <- summary(influence.measures(MainE_Model))

## VIF indicates we are fine
vif(MainE_Model)


## This can be added to the appendix as a sensitivity analysis 
## This is the final model
MainE_Model_subset <- lm(formula = log_losday ~ is30dayreadmit + cindex_cat + evisit + 
    ageyear + gender + maritalstatus + facilityname + insurancetype + 
    bpsystolic + temperature + heartrate + o2sat + respirationrate + 
    bpdiastolic, data = ghc_final, subset = -as.numeric(row.names(t)))

t2 <- MainE_Model_subset %>% broom::tidy() %>% kable(digits = 3)

```

There's `r  length(row.names(t))` influencial observations as flagged by different criterion. 

```{r validation_cv, cache = TRUE}

set.seed(1)

grand.MSE <- 0

folds <- c(rep(1,257),
           rep(2,257),
           rep(3,257),
           rep(4,257),
           rep(5,257),
           rep(6,257),
           rep(7,257),
           rep(8,257),
           rep(9,256),
           rep(10,256))

for (i in 1:100) {
  
  index <- sample(folds, size = 2568, replace = FALSE)
  
  MSE <- 0
  
  for (j in 1:10) {
    
    model <- lm(formula = log_losday ~ is30dayreadmit + cindex_cat + evisit + 
    ageyear + gender + maritalstatus + facilityname + insurancetype + 
    bpsystolic + temperature + heartrate + o2sat + respirationrate + 
    bpdiastolic, data = ghc_final, subset = which(index != j))
    
    MSE = c(MSE, mean((ghc_final$log_losday[which(index == j)] - 
                         predict(model, ghc_final[which(index == j), ]))^2))
    
  }
  
  grand.MSE <- c(grand.MSE, mean(MSE[-1]))
  
}

cv.MSE <- mean(grand.MSE[-1])
cv.MSE

mean((ghc_final$log_losday - predict(MainE_Model))^2)

```



```{r validation_boot, cache = T}

## Bootstrap 

boot.fn <- function(data, index){
	model <- lm(formula = log_losday ~ is30dayreadmit + evisit + ageyear + 
    gender + maritalstatus + facilityname + insurancetype + bpsystolic + 
    temperature + heartrate + respirationrate + bpdiastolic + 
    cindex_cat, data = data, subset = index)
	RSE <- summary(model)$sigma
	
	return(RSE^2)
	
}

boot.fn(ghc_final, 1:3258)

set.seed(1)

boot.output <- boot(ghc_final, boot.fn, 10000)

boot.ci(boot.output, type = "perc")

plot(boot.output)


```


